---
title: "Working with NOAA HPC Hera"
date: last-modified
date-format: iso
published-title: "Last updated"
author: Nicholas Ducharme-Barth
engine: knitr
execute:
    eval: false
    echo: true
    output: false
format:
  html:
    embed-resources: true
    toc: true
    toc-location: right
    number-sections: true
    code-overflow: wrap
---

## Example description & setup

In this example we step through submitting an array job on Hera where we want to run the same job in a number of directories. In this case the job is running a simple [R script](https://github.com/N-DucharmeBarth-NOAA/noaa-hpc-hera/blob/main/example_data/simple_r/script.r) that reads in the *.csv* file stored in the directory, fits a linear model, and writes the paramter estimates to a *.csv*. We specify which directories we want to run jobs in as a part of the job-array using a [text file](https://github.com/N-DucharmeBarth-NOAA/noaa-hpc-hera/blob/main/example_data/simple_r/simple_r_v2.txt) to specify the directory path names on Hera.

The *simple_r* example can be set-up either by cloning the repository `git clone https://github.com/N-DucharmeBarth-NOAA/noaa-hpc-hera.git`, or stepping through the following code:

Define relative path and create directory if needed

::: {.callout-important}
# Hard-coding alert!
Note you will have to change `proj_dir` to match your machine path and also may need to change `hera_project` from "NMFS/htc4sa/" if you are under a different HPC project.
:::

```{r}
proj_dir = "D:/HOME/SAP/Code/noaa-hpc-hera/"
dir.create(proj_dir, recursive=TRUE)
hera_project = "NMFS/htc4sa/Nicholas.Ducharme-barth/"
```

Define directory names
```{r}
dir_name = paste0("rep_0", 0:9)
```